# **Reproducible Research: Peer Assessment 1**


##### **_Loading the necessary libraries :_**
```{r} 
library(ggplot2)
library(dplyr)
```

## **Loading and preprocessing the data**
```{r}
activity_data <- read.csv("activity.csv")
head(activity_data)
```

##### **_Creating a histogram of number of steps taken each day :_**
```{r}
databydate <- activity_data %>% 
    select(date, steps) %>% 
    group_by(date) %>%
    summarise(tsteps=sum(steps)) %>%
    filter(!is.na(tsteps))

hist(databydate$tsteps, xlab="Total daily steps", 
     main="Frequency of total daily steps", breaks = 20)
```

### **What is mean and median total number of steps taken per day?**
```{r}
avg = mean(databydate$tsteps)
middle = median(databydate$tsteps)
```

* Mean : `r avg`
* Median : `r middle``


### **What is the average daily activity pattern?**
```{r}
databyinterval <- activity_data %>%
    select(interval, steps) %>% 
    filter(!is.na(steps)) %>%
    group_by(interval) %>%
    summarise(tsteps=mean(steps))

g <- ggplot(databyinterval, aes(x=interval, y=tsteps)) +
        geom_line() 
    
g + ggtitle("Time series of average number of steps taken") +
    theme(plot.title = element_text(hjust=0.5))
```

### **_The 5-minute interval that, on average, contains the maximum number of steps:_**
```{r}
max_steps <- databyinterval[which(databyinterval$tsteps == max(databyinterval$tsteps)),] 
```
* *maximum number of steps occurred at __`r max_steps$interval`__*

## **Imputing missing values**

##### **1. Calculating number of missing values**
```{r}
num_of_na_values <- sum(is.na(activity_data))
```

* *There are __`r num_of_na_values`__ missing values in the data-set.*

##### **2. Creating a new dataset by removing the missing values**
```{r}
new_data <- activity_data

for(i in 1:ncol(new_data)){
    new_data[is.na(new_data[,i]), i] <- mean(new_data[,i], na.rm = TRUE)
}

head(new_data)
```

##### ***3.Creating a histogram***
```{r}

new_databydate <- new_data %>% 
    select(date, steps) %>% 
    group_by(date) %>%
    summarise(tsteps=sum(steps))

hist(new_databydate$tsteps, xlab="Total daily steps", 
     main="Frequency of total daily steps", breaks = 20)
```

#### **4.Checking the mean and median of new_data**
```{r}
new_avg = mean(new_databydate$tsteps)
new_middle = median(new_databydate$tsteps)
```

- **Do these values differ from the estimates from the first part of the assignment? What is the impact of imputing missing data on the estimates of the total daily number of steps?**

- We know that 
  - New mean : `r new_avg` and Old mean : `r avg`
  - New median : `r new_middle` and Old median : `r middle`

So, the *median* has an impact on adding the missing values but not the *mean*.

### **Are there differences in activity patterns between weekdays and weekends?**
```{r}
new_data$date <- as.Date(new_data$date)
new_data$weekday <- weekdays(new_data$date)
new_data$weekend <- ifelse(new_data$weekday=="Saturday" | new_data$weekday=="Sunday", 
                           "Weekend", "Weekday")

averages <- aggregate(steps ~ interval + weekend, data =new_data, mean)

ggplot(averages, aes(x=interval, y=steps, color=weekend)) + 
    geom_line() + facet_grid(weekend ~ .) +
    labs(x="5-minute interval", y="Mean no of steps",
         title="Comparison of average number of steps in each interval")+
  theme(plot.title = element_text(hjust=0.5))

```

